---
title: "Assignment 5: Government Data API"
author: "Monica Campos"
---

```{r}
#| label: Scraping Government data Foreign Relations Committee
#| echo: true
## Scraping Government data
## Website: GovInfo (https://www.govinfo.gov/app/search/)
## Prerequisite: Download from website the list of files to be downloaded
## Designed for background job

# Start with a clean plate and lean loading to save memory

gc(reset=T)
library(purrr)
library(magrittr) # Alternatively, load tidyverse

## Set path for reading the listing and home directory
## For Windows, use "c:\\directory\\subdirectory\\"
## For Mac, "/Users/YOURNAME/path/"

setwd("/Users/monicacampos/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/Github Website")
library(rjson)
library(jsonlite)
library(data.table)
library(readr)

## JSON method
### rjson
gf_list <- rjson::fromJSON(file ="govinfo-search-results-2025-10-16T17_44_21.json")
govfile2=dplyr::bind_rows(gf_list$resultSet)


### Extract the list
govfiles3 <- gf_list$resultSet

### One more step
govfiles3 <- gf_list$resultSet |> dplyr::bind_rows()


# Preparing for bulk download of government documents
govfiles3$id = govfiles3$packageId
pdf_govfiles_url <- govfiles3$pdfLink
pdf_govfiles_id <- govfiles3$index

valid_indices <- !is.na(pdf_govfiles_url) & pdf_govfiles_url != ""
pdf_govfiles_url <- pdf_govfiles_url[valid_indices]
pdf_govfiles_id <- pdf_govfiles_id[valid_indices]

message(sprintf("Found %d valid PDF links to download", length(pdf_govfiles_url)))


# Directory to save the pdf's
# Be sure to create a folder for storing the pdf's
save_dir <- "/Users/monicacampos/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/Github Website"

# Function to download pdfs
download_govfiles_pdf <- function(url, id) {
  tryCatch({
    destfile <- paste0(save_dir, "govfiles_", id, ".pdf")
    download.file(url, destfile = destfile, mode = "wb") # Binary files
    Sys.sleep(runif(1, 1, 3))  # Important: random sleep between 1 and 3 seconds to avoid suspicion of "hacking" the server
    return(paste("Successfully downloaded:", url))
  },
  error = function(e) {
    return(paste("Failed to download:", url))
  })
}

# Download files, potentially in parallel for speed
# Simple timer, can use package like tictoc
# 

## Try downloading one document
start.time <- Sys.time()
message("Starting downloads")
results <- 1:1 %>% 
  purrr::map_chr(~ download_govfiles_pdf(pdf_govfiles_url[.], pdf_govfiles_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

## Try five
start.time <- Sys.time()
message("Starting downloads")
results <- 1:5 %>% 
  purrr::map_chr(~ download_govfiles_pdf(pdf_govfiles_url[.], pdf_govfiles_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken


# Try ten

start.time <- Sys.time()
message("Starting downloads")
tenresults <- 1:10 %>% 
  purrr::map_chr(~ download_govfiles_pdf(pdf_govfiles_url[.], pdf_govfiles_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

# Print results
print(tenresults)
```

Difficulties encountered in scrapping the data:

\(1\) Inefficient Download Process: I had sequential downloading which resulted in mandatory 1-3 seconds delay for approximately 25 minutes for 722 files. I did not have anything to previously download files to keep from redundant downloads and wasted bandwidth. When I ran the code, the same files were downloaded multiple times (file #1 downloaded 3 separate times)

\(2\) Limited Error Recovery: I do not know a way to track progress which makes it impossible to monitor long-running downloads. The absence of resume capability means interruptions require starting from scratch. Non-specific error message because it only shows "filed to download" without diagnostic details.

\(3\) File Management Issues: The file path concatenation error creates malformed file names that could overwrite each other. There needs to be further or more advanced file size validation. Finally, I had limited future analysis because the metadata from JSON source is not preserved alongside PDFs.

Data Usability Assessment: The overall usability is moderate, files are accessible, but there is still some limitations to be able to have more contextual information for systematic analysis.

-   Successfully retrieved 722 valid PDF links from official government source.

-   All test downloads (10/10) completed without errors.

-   Data source (govinfo.gov) is authoritative and reliable.

-   File naming convention captures unique identifiers for tracking.

Recommended Improvements:

1.  Fix file path constructing using `file.path()` to prevent naming errors.
2.  Implement existence checking to skip already-downloaded files.
3.  Export metadata CSV before downloading to preserve document context.
4.  Add progress bar using `progress` package for visibility.
5.  Enable parallel processing to reduce total by \~75%
6.  Create structured download log capturing success/failure status, time stamps, and error messages
7.  Implement resume functionality to continue from last successful downloads

```{r}
#| label: Exercise Foreign Affairs Committee Data Scraping
#| echo: true

# Start with a clean plate and lean loading to save memory

gc(reset=T)

library(purrr)
library(magrittr) # Alternatively, load tidyverse

## Set path for reading the listing and home directory
## For Windows, use "c:\\directory\\subdirectory\\"
## For Mac, "/Users/YOURNAME/path/"

setwd("/Users/monicacampos/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/Github Website")
library(rjson)
library(jsonlite)
library(data.table)
library(readr)

## JSON method
### rjson
fa_list <- rjson::fromJSON(file ="govinfo-search-results-2025-10-16T19_53_25.json")
govfilefa=dplyr::bind_rows(fa_list$resultSet)


### Extract the list
govfilefa <- fa_list$resultSet

### One more step
govfilefa <- fa_list$resultSet |> dplyr::bind_rows()


# Preparing for bulk download of government documents
govfilefa$id = govfilefa$packageId
pdf_govfilefa_url <- govfilefa$pdfLink
pdf_govfilefa_id <- govfilefa$index

valid_indices <- !is.na(pdf_govfiles_url) & pdf_govfiles_url != ""
pdf_govfilefa_url <- pdf_govfiles_url[valid_indices]
pdf_govfilefa_id <- pdf_govfiles_id[valid_indices]

message(sprintf("Found %d valid PDF links to download", length(pdf_govfiles_url)))


# Directory to save the pdf's
# Be sure to create a folder for storing the pdf's
save_dir <- "/Users/monicacampos/Library/CloudStorage/OneDrive-TheUniversityofTexasatDallas/Github Website"

# Function to download pdfs
download_govfilesfa_pdf <- function(url, id) {
  tryCatch({
    destfile <- paste0(save_dir, "govfiles_", id, ".pdf")
    download.file(url, destfile = destfile, mode = "wb") # Binary files
    Sys.sleep(runif(1, 1, 3))  # Important: random sleep between 1 and 3 seconds to avoid suspicion of "hacking" the server
    return(paste("Successfully downloaded:", url))
  },
  error = function(e) {
    return(paste("Failed to download:", url))
  })
}

# Download files, potentially in parallel for speed
# Simple timer, can use package like tictoc
# 

## Try downloading one document
start.time <- Sys.time()
message("Starting downloads")
results <- 1:1 %>% 
  purrr::map_chr(~ download_govfilesfa_pdf(pdf_govfilefa_url[.], pdf_govfilefa_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

## Try five
start.time <- Sys.time()
message("Starting downloads")
results <- 1:5 %>% 
  purrr::map_chr(~ download_govfilesfa_pdf(pdf_govfilefa_url[.], pdf_govfilefa_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken


# Try ten

start.time <- Sys.time()
message("Starting downloads")
results <- 1:10 %>% 
  purrr::map_chr(~ download_govfilesfa_pdf(pdf_govfilefa_url[.], pdf_govfilefa_id[.]))
message("Finished downloads")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

# Print results
print(tenresults)
```
