---
title: "Text Analytics Using Quanteda"
author: "Monica Campos"
---

a\. Biden-Xi summit data

```{r}
#| label: Biden-Xi Summit
#| echo: true
# Sample program for using quanteda for text modeling and analysis
# Documentation: vignette("quickstart", package = "quanteda")
# Website: https://quanteda.io/

library(quanteda)
library(quanteda.textmodels)
library(quanteda.textplots)
library(readr)
library(ggplot2)

# Twitter data about President Biden and Xi summit in Novemeber 2021
# Do some background search/study on the event
# 
summit <- read_csv("https://raw.githubusercontent.com/datageneration/datamethods/master/textanalytics/summit_11162021.csv")
View(summit)

sum_twt = summit$text
toks = tokens(sum_twt)
sumtwtdfm <- dfm(toks)
class(toks)

 # Latent Semantic Analysis 
## (https://quanteda.io/reference/textmodel_lsa.html)

sum_lsa <- textmodel_lsa(sumtwtdfm, nd=4,  margin = c("both", "documents", "features"))
summary(sum_lsa)
head(sum_lsa$docs)
class(sum_lsa)

tweet_dfm <- tokens(sum_twt, remove_punct = TRUE) %>%
  dfm()
head(tweet_dfm)

tag_dfm <- dfm_select(tweet_dfm, pattern = "#*")
toptag <- names(topfeatures(tag_dfm, 50))
head(toptag, 10)
library("quanteda.textplots")
tag_fcm <- fcm(tag_dfm)
head(tag_fcm)
topgat_fcm <- fcm_select(tag_fcm, pattern = toptag)
textplot_network(topgat_fcm, min_freq = 50, edge_alpha = 0.8, edge_size = 1)
user_dfm <- dfm_select(tweet_dfm, pattern = "@*")
topuser <- names(topfeatures(user_dfm, 50))
head(topuser, 20)
user_fcm <- fcm(user_dfm)
head(user_fcm, 20)

user_fcm <- fcm_select(user_fcm, pattern = topuser)
textplot_network(user_fcm, min_freq = 20, edge_color = "firebrick", edge_alpha = 0.8, edge_size = 1)
```

The analysis of the Biden-Xi Summit Twitter data, discusses the meeting between the two presidents at the November 2021 summit, contained 14,220 rows, 90 columns. The text analysis was made using the Algorithmic Modeling Culture for pattern discovery using R tools of quanteda and ggplot2 to demonstrate nd plot the process of turning noisy social media data into quantifiable structure to draw insights on nature of conversation.

There are three key findings that provide structural connectivity in a heavy political discussion:

\(1\) The corpus had 58,080 documents or tweets and 63, 972 features or unique words, which explains the following the needed dimension reduction that was successfully completed by Latent Semantic Analysis (LSA). The LSA reduced the complexity to four latent dimensions/topics. This means the Biden-Xi Summit can be summarized in four underlying themes.

\(2\) The second finding is the Feature Co-occurrence Matrix (FCM) of the hashtags that were used to revealed a strong, quantifiable relationship between entities. The highest co-occurrence of 443 times was between #biden and #china. Similarly, the hashtags between country and leader: #china and #xijinping had a strong linkage.

\(3\) Influence and User Context: The three official accounts (\@potus and \@joebiden), and major media and news sources (\@politico, \@nytimes) are identified as the top entities in the conversation. The final piece of the code was the creation of a matrix (user_fm) to map which users were mentioned together to draw insights and understanding on the flow of political discourse.

b\. US presidential inaugural speeches

```{r}
#| label: Inaugural Speeches
#| echo: true
# Sample program for using quanteda for text modeling and analysis
# Documentation: vignette("quickstart", package = c("quanteda.textstats", "quanteda.textmodels")
# Website: https://quanteda.io/

library(quanteda)
library(quanteda.textmodels)
library(quanteda.textplots)
library(readr)
library(ggplot2)

# Wordcloud
# based on US presidential inaugural address texts, and metadata (for the corpus), from 1789 to present.
dfm_inaug <- corpus_subset(data_corpus_inaugural, Year <= 1826) %>% 
  tokens(remove_punct = TRUE) %>% 
  tokens_remove(stopwords('english')) %>% 
  dfm() %>%
  dfm_trim(min_termfreq = 10, verbose = FALSE)

set.seed(100)
textplot_wordcloud(dfm_inaug)
inaug_speech = data_corpus_inaugural

corpus_subset(data_corpus_inaugural, 
              President %in% c("Trump", "Obama", "Bush")) %>%
  tokens(remove_punct = TRUE) %>%
  tokens_remove(stopwords("english")) %>%
  dfm() %>%
  dfm_group(groups = President) %>%
  dfm_trim(min_termfreq = 5, verbose = FALSE) %>%
  textplot_wordcloud(comparison = TRUE)


textplot_wordcloud(dfm_inaug, min_count = 10,
                   color = c('red', 'pink', 'green', 'purple', 'orange', 'blue'))


data_corpus_inaugural_subset <- 
  corpus_subset(data_corpus_inaugural, Year > 1949)
kwic(tokens(data_corpus_inaugural_subset), pattern = "american") %>%
  textplot_xray()


textplot_xray(
  kwic(tokens(data_corpus_inaugural_subset), pattern = "american"),
  kwic(tokens(data_corpus_inaugural_subset), pattern = "people"),
  kwic(tokens(data_corpus_inaugural_subset), pattern = "communist")
  
)

## Why is the "communist" plot missing?

theme_set(theme_bw())
g <- textplot_xray(
  kwic(tokens(data_corpus_inaugural_subset), pattern = "american"),
  kwic(tokens(data_corpus_inaugural_subset), pattern = "people"),
  kwic(tokens(data_corpus_inaugural_subset), pattern = "communist")
)
g + aes(color = keyword) + 
  scale_color_manual(values = c("blue", "red", "green")) +
  theme(legend.position = "none")


library(quanteda.textstats)
library(ggplot2)
features_dfm_inaug <- textstat_frequency(dfm_inaug, n = 100)

# Sort by reverse frequency order
features_dfm_inaug$feature <- with(features_dfm_inaug, reorder(feature, -frequency))

ggplot(features_dfm_inaug, aes(x = feature, y = frequency)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



# Get frequency grouped by president
freq_grouped <- textstat_frequency(dfm(tokens(data_corpus_inaugural_subset)), 
                                   groups = data_corpus_inaugural_subset$President)

# Filter the term "american"
freq_american <- subset(freq_grouped, freq_grouped$feature %in% "american")  

ggplot(freq_american, aes(x = group, y = frequency)) +
  geom_point() + 
  scale_y_continuous(limits = c(0, 14), breaks = c(seq(0, 14, 2))) +
  xlab(NULL) + 
  ylab("Frequency") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


dfm_rel_freq <- dfm_weight(dfm(tokens(data_corpus_inaugural_subset)), scheme = "prop") * 100
head(dfm_rel_freq)

rel_freq <- textstat_frequency(dfm_rel_freq, groups = dfm_rel_freq$President)

# Filter the term "american"
rel_freq_american <- subset(rel_freq, feature %in% "american")  

ggplot(rel_freq_american, aes(x = group, y = frequency)) +
  geom_point() + 
  scale_y_continuous(limits = c(0, 0.7), breaks = c(seq(0, 0.7, 0.1))) +
  xlab(NULL) + 
  ylab("Relative frequency") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


dfm_weight_pres <- data_corpus_inaugural %>%
  corpus_subset(Year > 2000) %>%
  tokens(remove_punct = TRUE) %>%
  tokens_remove(stopwords("english")) %>%
  dfm() %>%
  dfm_weight(scheme = "prop")

# Calculate relative frequency by president
freq_weight <- textstat_frequency(dfm_weight_pres, n = 10, 
                                  groups = dfm_weight_pres$President)

ggplot(data = freq_weight, aes(x = nrow(freq_weight):1, y = frequency)) +
  geom_point() +
  facet_wrap(~ group, scales = "free") +
  coord_flip() +
  scale_x_continuous(breaks = nrow(freq_weight):1,
                     labels = freq_weight$feature) +
  labs(x = NULL, y = "Relative frequency")


# Only select speeches by Obama and Trump
pres_corpus <- corpus_subset(data_corpus_inaugural, 
                             President %in% c("Obama", "Trump"))

# Create a dfm grouped by president
pres_dfm <- tokens(pres_corpus, remove_punct = TRUE) %>%
  tokens_remove(stopwords("english")) %>%
  tokens_group(groups = President) %>%
  dfm()

# Calculate keyness and determine Trump as target group
result_keyness <- textstat_keyness(pres_dfm, target = "Trump")

# Plot estimated word keyness
textplot_keyness(result_keyness) 

# Plot without the reference text (in this case Obama)
textplot_keyness(result_keyness, show_reference = FALSE)


library(quanteda.textmodels)

# Irish budget speeches from 2010 (data from quanteda.textmodels)
# Transform corpus to dfm
data(data_corpus_irishbudget2010, package = "quanteda.textmodels")
ie_dfm <- dfm(tokens(data_corpus_irishbudget2010))

# Set reference scores
refscores <- c(rep(NA, 4), 1, -1, rep(NA, 8))

# Predict Wordscores model
ws <- textmodel_wordscores(ie_dfm, y = refscores, smooth = 1)

# Plot estimated word positions (highlight words and print them in red)
textplot_scale1d(ws,
                 highlighted = c("minister", "have", "our", "budget"), 
                 highlighted_color = "red")


# Get predictions
pred <- predict(ws, se.fit = TRUE)

# Plot estimated document positions and group by "party" variable
textplot_scale1d(pred, margin = "documents",
                 groups = docvars(data_corpus_irishbudget2010, "party"))

# Plot estimated document positions using the LBG transformation and group by "party" variable

pred_lbg <- predict(ws, se.fit = TRUE, rescaling = "lbg")

textplot_scale1d(pred_lbg, margin = "documents",
                 groups = docvars(data_corpus_irishbudget2010, "party"))


# Estimate Wordfish model
library("quanteda.textmodels")
wf <- textmodel_wordfish(dfm(tokens(data_corpus_irishbudget2010)), dir = c(6, 5))

# Plot estimated word positions
textplot_scale1d(wf, margin = "features", 
                 highlighted = c("government", "global", "children", 
                                 "bank", "economy", "the", "citizenship",
                                 "productivity", "deficit"), 
                 highlighted_color = "red")


# Plot estimated document positions
textplot_scale1d(wf, groups = data_corpus_irishbudget2010$party)


# Transform corpus to dfm
ie_dfm <- dfm(tokens(data_corpus_irishbudget2010))

# Run correspondence analysis on dfm
ca <- textmodel_ca(ie_dfm)

# Plot estimated positions and group by party
textplot_scale1d(ca, margin = "documents",
                 groups = docvars(data_corpus_irishbudget2010, "party"))
```

This code analyzes the United States Presidential Inaugural Speeches using a variety of techniques (word clouds, KWIC/X-Ray Plots, frequency measures, and Keyness analysis) to track rhetorical changes over time and stylistic differences between leaders.

i.  Any similarities and differences over time and among presidents?

Presidents became less likely to refer to the country as the "Union" and less likely to mention "Constitution" or "Foreign" countries. As time passed they were more likely to refer to the United States as "America" and more likely to mention "freedom" and "God."

c\. What is Wordfish? 

Based on the quanteda website, Wordfish is an unsupervised scaling model used to estimate the relative ideological or policy positions of texts/documents with a single, latent dimensions. Also, it is considered a Poisson scaling model that uses the maximum likelihood estimation to simultaneously determine both the estimated positions of the document and words (features) within the text. Wordfish does not require any assignment of policy position or reference scores. The only thing it needs is to set the orientation and document of the scale using two specific docuents (e.g. dir = c(6,5))).

Learn more about WordFish: <https://quanteda.io/articles/pkgdown/examples/plotting.html?q=wordfish#wordfish>)
